Is partner choice related to prosociality? A cross-national investigation

### 2022-09-23 manuscript.Rmd
resubmission

Here, we report the results of two pre-registered studies of the cross-national associations between relational mobility, our socioecological proxy for partner choice, and a range of prosocial behaviours and attitudes. In Study 1, we leveraged data from the Global Preferences Survey [@Falk2018], a cross-national study of social preferences including impersonal altruism, positive reciprocity, and generalised trust. We focused on these particular measures of prosociality because altruistic, reciprocal, and trusting behaviours have been shown to reflect a single behavioural construct dubbed the "cooperative phenotype" in previous work [@Peysakhovich2014]. All three of these behaviours are predicted to increase under higher levels of partner choice: altruistic and reciprocal prosocial behaviours become useful as signals of cooperative intent for potential partners, and levels of trust thus increase along with levels of prosociality in the population. In Study 2, we used variables from the World Values Survey [@WVS2014] measuring collective action, moral assessments of cheating behaviour, and trust, which additionally capture people's prosocial contribution to social dilemmas and willingness to uphold prosocial moral norms.

@@

The Global Preferences Survey was conducted as part of the 2012 World Gallup Poll (https://www.gallup.com/analytics/213704/world-poll.aspx). The World Gallup Poll is conducted either via telephone or via face-to-face interview. For telephone interviews, nationally representative samples were achieved through the use of random-digit dialling or nationally representative lists of phone numbers. For face-to-face interviews, nationally representative samples were achieved through the use of a random route procedure within primary sampling units stratified by geography and/or population size.

@@

Second, we controlled for geographic and linguistic proximity between countries. Countries that are close to one another and share common cultural ancestors are likely to be more similar to one another, due to similar ecologies, climates, institutions, and norms (see Figure \@ref(fig:plotDAG)). To account for these unmeasured confounds, we allowed countries to covary according to geographic and linguistic proximity in our models. Geographic proximity was calculated as the inverse of the logged geodesic distance between country capital cities [data from the R package _maps_, @Brownrigg2018] using the R package _geosphere_ [@Hijmans2019]. Linguistic proximity between two countries was calculated as the cultural proximity between all languages spoken within those countries, weighted by speaker percentages [@Glottolog; @Ethnologue]: see Supplementary Methods for more details. While we acknowledge that linguistic proximity might not capture all cultural similarities between nations (e.g., differences in social structure and institutions between majority Spanish-speaking countries), this proxy for cultural relatedness has been shown to predict a number of national-level outcomes around the world, such as democratic institutions [@Matthews2016] and human development [@Claessens_preprint].

>>>>

Here, we report the results of two pre-registered studies of the cross-national associations between relational mobility, our socioecological proxy for partner choice, and a range of prosocial behaviours and attitudes. In Study 1, we leveraged data from the Global Preferences Survey [@Falk2018], a cross-national study of social preferences including impersonal altruism, positive reciprocity, and generalised trust. We focused on these particular measures of prosociality because altruistic, reciprocal, and trusting behaviours have been shown to reflect a single behavioural construct dubbed the "cooperative phenotype" in previous work [@Peysakhovich2014]. All three of these behaviours are predicted to increase under higher levels of partner choice: altruistic and reciprocal prosocial behaviours become useful as signals of cooperative intent for potential partners, especially when broadcasted publicly, and levels of trust thus increase along with levels of prosociality in the population. In Study 2, we used variables from the World Values Survey [@WVS2014] measuring collective action, moral assessments of cheating behaviour, and trust, which additionally capture people's prosocial contribution to social dilemmas and willingness to uphold prosocial moral norms.

@@

The Global Preferences Survey was conducted as part of the 2012 Gallup World Poll (https://www.gallup.com/analytics/318875/global-research.aspx). The Gallup World Poll is conducted either via telephone or via face-to-face interview. For telephone interviews, nationally representative samples were achieved through the use of random-digit dialling or nationally representative lists of phone numbers. For face-to-face interviews, nationally representative samples were achieved through the use of a random route procedure within primary sampling units stratified by geography and/or population size.

@@

Second, we controlled for geographic and linguistic proximity between countries. Countries that are close to one another and share common cultural ancestors are likely to be more similar to one another, due to similar ecologies, climates, institutions, and norms (see Figure \@ref(fig:plotDAG)). To account for these unmeasured confounds, we allowed countries to covary according to geographic and linguistic proximity in our models. Geographic proximity was calculated as the inverse of the logged geodesic distance between country capital cities [data from the R package _maps_, @Brownrigg2018] using the R package _geosphere_ [@Hijmans2019]. Linguistic proximity between two countries was calculated as the cultural proximity between all languages spoken within those countries, weighted by speaker percentages [@Glottolog; @Ethnologue]: see Supplementary Methods for more details.

### 2022-09-23 manuscript.tex
resubmission

Here, we report the results of two pre-registered studies of the cross-national associations between relational mobility, our socioecological proxy for partner choice, and a range of prosocial behaviours and attitudes. In Study 1, we leveraged data from the Global Preferences Survey (Falk et al., 2018), a cross-national study of social preferences including impersonal altruism, positive reciprocity, and generalised trust. We focused on these particular measures of prosociality because altruistic, reciprocal, and trusting behaviours have been shown to reflect a single behavioural construct dubbed the ``cooperative phenotype'' in previous work (Peysakhovich et al., 2014). All three of these behaviours are predicted to increase under higher levels of partner choice: altruistic and reciprocal prosocial behaviours become useful as signals of cooperative intent for potential partners, and levels of trust thus increase along with levels of prosociality in the population. In Study 2, we used variables from the World Values Survey (Inglehart et al., 2014) measuring collective action, moral assessments of cheating behaviour, and trust, which additionally capture people's prosocial contribution to social dilemmas and willingness to uphold prosocial moral norms.

@@

The Global Preferences Survey was conducted as part of the 2012 World Gallup Poll (\url{https://www.gallup.com/analytics/213704/world-poll.aspx}). The World Gallup Poll is conducted either via telephone or via face-to-face interview. For telephone interviews, nationally representative samples were achieved through the use of random-digit dialling or nationally representative lists of phone numbers. For face-to-face interviews, nationally representative samples were achieved through the use of a random route procedure within primary sampling units stratified by geography and/or population size.

@@

Second, we controlled for geographic and linguistic proximity between countries. Countries that are close to one another and share common cultural ancestors are likely to be more similar to one another, due to similar ecologies, climates, institutions, and norms (see Figure \ref{fig:plotDAG}). To account for these unmeasured confounds, we allowed countries to covary according to geographic and linguistic proximity in our models. Geographic proximity was calculated as the inverse of the logged geodesic distance between country capital cities (data from the R package \emph{maps}, Brownrigg, 2018) using the R package \emph{geosphere} (Hijmans, 2019). Linguistic proximity between two countries was calculated as the cultural proximity between all languages spoken within those countries, weighted by speaker percentages (Eberhard et al., 2018; Hammarström et al., 2017): see Supplementary Methods for more details. While we acknowledge that linguistic proximity might not capture all cultural similarities between nations (e.g., differences in social structure and institutions between majority Spanish-speaking countries), this proxy for cultural relatedness has been shown to predict a number of national-level outcomes around the world, such as democratic institutions (Matthews et al., 2016) and human development (Claessens \& Atkinson, 2022).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Matthews2016}{}}%
Matthews, L. J., Passmore, S., Richard, P. M., Gray, R. D., \& Atkinson, Q. D. (2016). Shared cultural history as a predictor of political and economic changes among nation states. \emph{PLOS ONE}, \emph{11}(4), 1--18. \url{https://doi.org/10.1371/journal.pone.0152979}

>>>>

Here, we report the results of two pre-registered studies of the cross-national associations between relational mobility, our socioecological proxy for partner choice, and a range of prosocial behaviours and attitudes. In Study 1, we leveraged data from the Global Preferences Survey (Falk et al., 2018), a cross-national study of social preferences including impersonal altruism, positive reciprocity, and generalised trust. We focused on these particular measures of prosociality because altruistic, reciprocal, and trusting behaviours have been shown to reflect a single behavioural construct dubbed the ``cooperative phenotype'' in previous work (Peysakhovich et al., 2014). All three of these behaviours are predicted to increase under higher levels of partner choice: altruistic and reciprocal prosocial behaviours become useful as signals of cooperative intent for potential partners, especially when broadcasted publicly, and levels of trust thus increase along with levels of prosociality in the population. In Study 2, we used variables from the World Values Survey (Inglehart et al., 2014) measuring collective action, moral assessments of cheating behaviour, and trust, which additionally capture people's prosocial contribution to social dilemmas and willingness to uphold prosocial moral norms.

@@

The Global Preferences Survey was conducted as part of the 2012 Gallup World Poll (\url{https://www.gallup.com/analytics/318875/global-research.aspx}). The Gallup World Poll is conducted either via telephone or via face-to-face interview. For telephone interviews, nationally representative samples were achieved through the use of random-digit dialling or nationally representative lists of phone numbers. For face-to-face interviews, nationally representative samples were achieved through the use of a random route procedure within primary sampling units stratified by geography and/or population size.

@@

Second, we controlled for geographic and linguistic proximity between countries. Countries that are close to one another and share common cultural ancestors are likely to be more similar to one another, due to similar ecologies, climates, institutions, and norms (see Figure \ref{fig:plotDAG}). To account for these unmeasured confounds, we allowed countries to covary according to geographic and linguistic proximity in our models. Geographic proximity was calculated as the inverse of the logged geodesic distance between country capital cities (data from the R package \emph{maps}, Brownrigg, 2018) using the R package \emph{geosphere} (Hijmans, 2019). Linguistic proximity between two countries was calculated as the cultural proximity between all languages spoken within those countries, weighted by speaker percentages (Eberhard et al., 2018; Hammarström et al., 2017): see Supplementary Methods for more details.

@@
